\birthplace{Shanghai}
\birthday{October 12th 1983}

\school{University of Oregon, Eugene, OR}
\school{Shanghai Jiao Tong University, Shanghai, China}

\degree{Master of Science in Computer Science, 2008,
University of Oregon}
\degree{Bachelor of Information Security, 2006,
Shanghai Jiao Tong University.}
\interests{Data mining, machine learning, the Semantic Web}

%\position{Graduate Research \& Teaching Assistant, Department of Computer and Information Science, University of Oregon, 2006 to present}
%\position{Research and Development Intern, Pengyue Technology, Shanghai, 2006}

\award{Graduate Teaching \& Research Fellowship, Computer and Information Science, 2006 to present}
\award{NSF Travel Grant, International Semantic Web Conference, 2010}
\award{NSF Travel Grant, International Conference on Data Mining, 2011}

\publication{Haishan Liu, Gwen Frishkoff, Robert Frank, Dejing Dou, Sharing and integration of cognitive neuroscience data: Metric and pattern matching across heterogeneous ERP datasets, Neurocomputing, Volume 92, 1 September 2012, Pages 156-169, ISSN 0925-2312.}

\publication{Haishan Liu, Paea LePendu, Ruoming Jin, and Dejing Dou. A Hypergraph-based Method for Discovering Semantically Associated Itemsets. In Proceedings of the 11th IEEE International Conference on Data Mining (ICDM), 2011.}

\publication{Haishan Liu and Dejing Dou. Breaking the Deadlock: Simultaneously Discovering Attribute Matching and Cluster Matching with Multi-Objective Simulated Annealing. In Proceedings of the International Conference on Ontologies, Databases and Application of Semantics (ODBASE), pages 698--715, 2011.}

\publication{Dejing Dou, Han Qin, and Haishan Liu. Semantic Translation for Rule-based Knowledge in Data Mining. In Proceedings of the 22nd International Conference on Database and Expert Systems Applications (DEXA), part II, pages 74--89, 2011.}

\publication{Haishan Liu. Towards Semantic Data Mining. In: Doctoral Consortium of the 9th International Semantic Web Conference (ISWC), November 2010.}

\publication{Christopher Townsend, Jingshan Huang, Dejing Dou, Shivraj Dalvi, Patrick J. Hayes, Lei He, Wen-chang Lin, Haishan Liu, Robert Rudnick and Hardik Shah. OMIT: Domain Ontology and Knowledge Acquisition in MicroRNA Target Prediction. On-The-Move Conferences 2010 (ODBASE): 1160--1167}

\publication{Haishan Liu, Gwen Frishkoff, Robert Frank, and Dejing Dou. Ontology-based Mining of Brainwaves: A Sequence Similarity Technique for Mapping Alternative Descriptions of Patterns in Event Related Potentials (ERP) Data. In Proceedings of the 14th Pacific-Asia Conference on Knowledge Discovery and Data Mining (PAKDD), pages 43--54, 2010.}

\publication{Gwen Frishkoff, Paea LePendu, Robert Frank, Haishan Liu, and Dejing Dou. Development of Neural Electromagnetic Ontologies (NEMO): Ontology-based Tools for Representation and Integration of Event-related Brain Potentials. In Proceedings of the International Conference on Biomedical Ontology (ICBO), pages 31--34, 2009.}

%\publication{Haishan Liu, Gwen Frishkoff, Robert Frank and Dejing Dou. Mapping ERP patterns from heterogeneous datasets: A simulation study. Poster in the 15th Annual Meeting of the Organization for Human Brain Mapping (HBM). June, 2009.}

\publication{Haishan Liu and Dejing Dou. An Exploration of Understanding Heterogeneity through Data Mining. In Proceedings of KDD'08 Workshop on Mining Multiple Information Sources, pages 18--25, 2008.}

\publication{Jongwan Kim, Dejing Dou, Haishan Liu, and Donghwi Kwak. Constructing a user preference ontology for anti-spam mail systems. In Proceedings of the 20th Canadian Conference on Artificial Intelligence (Canadian AI), pages 272--283, 2007.} 